The Rise and Fall of the Note: Changing Paper Lengths in ACM CSCW, 2000--2018

### 2019-08-28 paper/paper.tex
add DOI, update copyright

%
% These commands are for a JOURNAL article.
\setcopyright{acmcopyright}
\acmJournal{PACMHCI}
\acmYear{2019}\acmVolume{X}\acmNumber{X}\acmArticle{X}\acmMonth{11}
\acmDOI{10.1145/XXXXXX.XXXXXX}

@@

The second major change in ACM CSCW publishing covered is around the shift to a new single-column document format, with substantially fewer words per page. This format shift was introduced alongside publishing papers in the \textit{Proceedings of the ACM on Human-Computer Interaction} (PACMHCI), and from holding CSCW in the Spring to the Fall. Given the 18 month gap between CSCW 2017 and 2018, two rounds of submissions were held for CSCW 2018. The first was published in 2017 an ``Online First'' edition of PACMHCI, labeled as ``2017.5.'' In addition, a change in the ACM reference format starting in 2016 led to longer character lengths per reference, which is not examined in this note, but is controlled for. CSCW also moved from a bi-annual to an annual conference in 2011. Because of the simultaneous changes in the CSCW community and in research and academia more generally over this period, I hesitate to draw strict causal claims and consider this note an exploratory work. However, this dataset and these conditions are as an excellent opportunity to investigate the continued development of the genre of the CSCW paper.

@@

With the raw text files, I determined the start of reference and appendix sections using regular expressions to search for various phrasings and spellings of these headings. I looked for the last occurrence of one of the phrases on a line either by itself or surrounded by any number of non-alphabetical characters. \footnote{References: \texttt{\textasciicircum{}[\textasciicircum{}a-zA-Z]*(bibliography|references|reference|works cited|refefences)[\textasciicircum{}a-zA-Z]*\$} \newline
Appendices: \texttt{\textasciicircum{}[\textasciicircum{}a-zA-Z]*(appendix|appendices|appendixes)[\textasciicircum{}a-zA-Z]*\$} } I extracted the text of the reference section and appendices, based on which section occurred first, as the order varied. I manually examined papers that were extreme outliers in terms of reference section length or where no reference sections were found. I found 15 such papers, which were errors in document processing that were manually corrected in Word .docx format, then converted to plain text in the same pandoc pipeline. I then computed the number of characters, words, and references in reference sections, and words were defined as tokens separated by any number of contiguous spaces. Data analysis used the pandas \cite{pandas} library, and visualizations were made using the matplotlib \cite{matplotlib} and seaborn \cite{seaborn} libraries. Analyses were conducted in Jupyter Notebooks \cite{jupyter}, which are publicly available\footnote{\url{https://github.com/staeiou/cscw19-paper-lengths}} and can be run on Binder \cite{binder} --- although the paper text is copyrighted and not publicly released.

>>>>

%
% These commands are for a JOURNAL article.
\setcopyright{rightsretainedccby}
\acmJournal{PACMHCI}
\acmYear{2019} \acmVolume{3} \acmNumber{CSCW} \acmArticle{222} \acmMonth{11} \acmPrice{15.00}\acmDOI{10.1145/3359324}

@@

The second major change in ACM CSCW publishing covered is around the shift to a new single-column document format, with substantially fewer words per page. This format shift was introduced alongside publishing papers in the \textit{Proceedings of the ACM on Human-Computer Interaction} (PACMHCI), and from holding CSCW in the Spring to the Fall. Given the 18 month gap between CSCW 2017 and 2018, two rounds of submissions were held for CSCW 2018. The first was published in a 2017 ``Online First'' edition of PACMHCI, labeled as ``2017.5.'' In addition, a change in the ACM reference format starting in 2016 led to longer character lengths per reference, which is not examined in this note, but is controlled for. CSCW also moved from a bi-annual to an annual conference in 2011. Because of the simultaneous changes in the CSCW community and in research and academia more generally over this period, I hesitate to draw strict causal claims and consider this note an exploratory work. However, this dataset and these conditions are as an excellent opportunity to investigate the continued development of the genre of the CSCW paper.

@@

With the raw text files, I determined the start of reference and appendix sections using regular expressions to search for various phrasings and spellings of these headings. I looked for the last occurrence of one of the phrases on a line either by itself or surrounded by any number of non-alphabetical characters. \footnote{References: \texttt{\textasciicircum{}[\textasciicircum{}a-zA-Z]*(bibliography|references|reference|works cited|refefences)[\textasciicircum{}a-zA-Z]*\$} \newline
Appendices: \texttt{\textasciicircum{}[\textasciicircum{}a-zA-Z]*(appendix|appendices|appendixes)[\textasciicircum{}a-zA-Z]*\$} } I extracted the text of the reference section and appendices, based on which section occurred first, as the order varied. I manually examined papers that were extreme outliers in terms of reference section length or where no reference sections were found. I found 15 such papers, which were errors in document processing that were manually corrected in Word .docx format, then converted to plain text in the same pandoc pipeline. I then computed the number of characters, words, and references in reference sections, and words were defined as tokens separated by any number of contiguous spaces. Data analysis used the pandas \cite{pandas} library, and visualizations were made using the matplotlib \cite{matplotlib} and seaborn \cite{seaborn} libraries. Analyses were conducted in Jupyter Notebooks \cite{jupyter}, which are publicly available\footnote{GitHub: \url{https://github.com/staeiou/cscw19-paper-lengths}; Zenodo: \url{https://doi.org/10.5281/zenodo.3380345}} and can be run on Binder \cite{binder} --- although the paper text is copyrighted and not publicly released.
